Estimated fitness effects of amino-acid mutations to SARS-CoV-2 proteins

### 2023-01-14 paper/paper.tex
add references and small edits to paper.tex (#12)
Co-authored-by: Richard Neher <richard.neher@unibas.ch>

\newcommand\jdbcomment[1]{\textcolor{red}{[#1]}}

@@

An alternative is sequence-based estimates, which so far have focused largely on clade growth. \jdbcomment{Richard, can you note any key papers you think should be cited as background.}

Here we describe new approach based on comparing expected versus observed counts.

>>>>

\newcommand\jdbcomment[1]{\textcolor{red}{[#1]}}
\newcommand\rancomment[1]{\textcolor{blue}{[#1]}}

@@

An alternative is sequence-based estimates, which so far have focused largely on clade growth and identification of mutations that mediate immune escape or increase transmissibility \citep{obermeyer2022analysis,lee2022inferring}.
Mutations that are repeatedly observed at the base of clades that grow faster than the background likely contribute to growth.
Such adaptive mutations are only a small minority of all available mutations and which mutations are adaptive depends on the environment (immunity) and the genetic background.
The great majority of mutations, however, have either small fitness effects (approximately neutral) or are deleterious to viral replication.
Here, we describe an approach to quantify the fitness cost of these mutations based on comparing expected versus observed counts.

\rancomment{older paper on flu where not counts but trajectories were used to partition sequences into good and bad \citep{strelkowa2021clonal}}
